### 2017-01-19 Projects/DelCo_2014_2015/Reports/Propane_Gas_Fire_Experiments/TeX_Files/Propane_Gas_Fire_Experiments.tex
Added Nelson's edits to gas burner report

\subsubsection{Leakage}
An air leakage measurement system was used to measure the amount of leakage associated with each structure. The amount of leakage in the East Structure was measured as 0.024~m$^2$. For the West Structure, the leakage was measured as 0.027~m$^2$ when the stairway door was fully closed, 0.054~m$^2$ when the stairway door was fully opened, and 0.048~m$^2$ when the stairway door was in the ``closed'' position (having a 152~mm (6~in) gap between the door and the frame) used Tests~24 and 25.
\clearpage

@@

\section{Instrumentation}
\label{sec:Instrumentation}
The structures were instrumented for temperature, gas velocity, heat flux, and gas concentration measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with a 3.2~mm (0.13~in) diameter inconel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. Calibrated pumps pulled gas samples through a sample conditioning system to eliminate moisture in the sample. Then, the dry gas samples were piped to a series of gas analyzers and the concentrations of oxygen, carbon monoxide, and carbon dioxide were measured. A legend is presented in Fig.~\ref{fig:Instrumentation_Legend} to clarify the instrumentation schematics presented in the follow sections.

@@

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}
This section lists the uncertainties in the reported length, mass, temperature, heat flux, gas concentration, gas velocity, and heat release rate measurements. Uncertainty estimates are based either on manufacturer literature or analyses performed by others for similar measurement devices and techniques. In accordance with NIST guidelines~\cite{Taylor&Kuyatt:1994}, measurement accuracy is reported as an {\em expanded uncertainty} or 95~\% ($2\,\sigma$) confidence interval. Most manufactures express accuracy in terms of a {\em standard  uncertainty} or 68~\% ($1\,\sigma$) confidence interval.

\subsubsection{Compartment Dimensions}
Room dimensions and instrumentation array locations were made with a hand held laser measurement device with a standard uncertainty of $\pm$6.0~mm (0.25~in) over a range of 0.6~m (2.0~ft) to 15~m (50.0~ft)~\cite{StanleyTools}. Steel measuring tapes with a resolution of $\pm$0.5~mm (0.02~in) were used to locate measurement devices. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm$1.6~mm (0.06~in) for 9.1~m (30~ft) tapes and $\pm$6.4~mm (0.25~in) for 30.5~m (100~ft) tapes~\cite{Butcher:2012}. These uncertainties are all well within the precision of the reported dimensions, which are typically rounded to the nearest 10~cm (4~in).

\subsubsection{Thermocouples}
The standard uncertainty in the temperature of the thermocouple wire itself as determined by the wire manufacturer, OMEGA Engineering, Inc., is $\pm$2.2$~^{\circ}$C at $277~^{\circ}$C and increases to $\pm$9.5$~^{\circ}$C at $871~^{\circ}$C~\cite{Omega:2004}. The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty. Expanded uncertainties as high as 20~\% for upper layer temperatures measured by a 1~mm bare-bead type K thermocouple have been reported by NIST researchers~\cite{Blevins:1999,Pitts:2003}. Small diameter thermocouples were used during these experiments to limit the impact of radiative heating and cooling. The estimated relative expanded uncertainty associated with the temperature measurements is $\pm$15~\%.

\subsubsection{Heat Flux Gauges}
Total heat flux measurements were made with water-cooled Schmidt-Boelter gauges. The manufacturer, MEDTHERM Corporation, reports a $\pm$3~\% calibration expanded uncertainty for these devices~\cite{Medtherm:2003}. Results from an international study on total heat flux gauge calibration and response demonstrated that the relative expanded uncertainty of a Schmidt-Boelter gauge is typically $\pm$8~\%~\cite{Pitts:2006}.

\subsubsection{Gas Sampling}
A gas sampling system from California Analytical Instruments, Inc. (model 602P) having a relative expanded uncertainty of $\pm$1~\% when compared to span gas volume fractions~\cite{Bundy:2007} was used to make gas concentration measurements. However, according to a study by Lock~et~al.~\cite{Lock:1}, the non-uniformity and movement of exhaust gases contribute to an estimated relative expanded uncertainty of $\pm$12~\%.

\subsubsection{Bi-Directional Probes}
Bi-directional probes with Setra 264 pressure transducers from Setra Systems, Inc. were used to measure gas velocity in doorways. Bryant~\cite{Bryant:FSJ2009} estimates that the relative expanded uncertainty of this measurement ranges from $\pm$14~\% to $\pm$22~\%.

\subsubsection{Heat Release Rate}
A rotary type gas meter was used to measure the volume flow rate of propane into the gas burners. The manufacturer, Romet Limited, reports a relative standard uncertainty of $\pm$2~\%~\cite{Romet:2014} for this type of meter (model RM-3000). A volumetric flow rate was calculated from the gas meter volume readings and used in conjunction with the heat of combustion of propane to calculate the heat release rate of the fire. The total expanded uncertainty for the heat release rate obtained from this method is estimated to be $\pm8$~\%.

@@

\chapter{Summary}
\label{chap:Summary}
Nine full-scale fire tests were conducted in two residential-sized structures to examine how changing ventilation patterns within a structure affects the fire environment. The fire source for each experiment was provided by a set of three diffusion flame burners with propane as the fuel. Various doors and vents were opened and closed during each test to change ventilation within the structure. A PPV fan was also used during some of the experiments to change the ventilation patterns. Local measurements of temperature, gas velocity, heat flux, and gas concentrations were collected at various locations throughout the structure during the experiments. The total volume of propane delivered to the burners was measured by a rotary gas meter and was used to calculate the heat release rate of the fire during each test.

>>>>

\subsubsection{Leakage}
An air leakage measurement system was used to measure the leakage area associated with each structure. The leakage area in the East Structure was measured as 0.024~m$^2$. For the West Structure, the leakage was measured as 0.027~m$^2$ when the stairway door was fully closed, 0.054~m$^2$ when the stairway door was fully opened, and 0.048~m$^2$ when the stairway door was in the ``closed'' position (having a 152~mm (6~in) gap between the door and the frame) used Tests~24 and 25.
\clearpage

@@

\section{Instrumentation}
\label{sec:Instrumentation}
The structures were instrumented for temperature, gas velocity, heat flux, and gas concentration measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with a 3.2~mm (0.13~in) diameter inconel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium window to prevent contributions from convective heat transfer. Calibrated pumps pulled gas samples through a sample conditioning system to eliminate moisture in the sample. Then, the dry gas samples were piped to a series of gas analyzers and the concentrations of oxygen, carbon monoxide, and carbon dioxide were measured. A legend is presented in Fig.~\ref{fig:Instrumentation_Legend} to clarify the instrumentation schematics presented in the follow sections.

@@

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}
This section lists the uncertainties in the reported length, mass, temperature, heat flux, gas concentration, gas velocity, and heat release rate measurements. Uncertainty estimates are based either on manufacturer literature or analyses performed by others for similar measurement devices and techniques. In accordance with NIST guidelines~\cite{Taylor&Kuyatt:1994}, measurement accuracy is reported as an {\em expanded uncertainty} or 95~\% ($2\,\sigma$) confidence interval. Most manufacturer specifications express accuracy in terms of a {\em standard  uncertainty} or 68~\% ($1\,\sigma$) confidence interval.

\subsubsection{Compartment Dimensions}
Room dimensions and instrumentation array locations were made with a hand held laser measurement device with a standard uncertainty of $\pm$6.0~mm (0.25~in) over a range of 0.6~m (2.0~ft) to 15~m (50.0~ft) according to the manufacturer~\cite{StanleyTools}. Steel measuring tapes with a resolution of $\pm$0.5~mm (0.02~in) were used to locate measurement devices. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm$1.6~mm (0.06~in) for 9.1~m (30~ft) tapes and $\pm$6.4~mm (0.25~in) for 30.5~m (100~ft) tapes~\cite{Butcher:2012}. These uncertainties are all well within the precision of the reported dimensions, which are typically rounded to the nearest 10~cm (4~in).

\subsubsection{Thermocouples}
The standard uncertainty in the temperature of the thermocouple wire itself as determined by the wire manufacturer, OMEGA Engineering, Inc., is $\pm$2.2$~^{\circ}$C at $277~^{\circ}$C and increases to $\pm$9.5$~^{\circ}$C at $871~^{\circ}$C~\cite{Omega:2004}. The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty. Expanded uncertainties as high as 20~\% for upper layer temperatures measured by a 1~mm bare-bead type K thermocouple have been reported by NIST researchers~\cite{Blevins:1999,Pitts:2003}. Small diameter thermocouples were used during these experiments to limit the impact of radiative heating and cooling. The estimated expanded uncertainty associated with the temperature measurements is $\pm$15~\%.

\subsubsection{Heat Flux Gauges}
Total heat flux measurements were made with water-cooled Schmidt-Boelter gauges. The manufacturer, MEDTHERM Corporation, reports a $\pm$3~\% calibration expanded uncertainty for these devices~\cite{Medtherm:2003}. Results from an international study on total heat flux gauge calibration and response demonstrated that the expanded uncertainty of a Schmidt-Boelter gauge is typically $\pm$8~\%~\cite{Pitts:2006}.

\subsubsection{Gas Sampling}
A gas sampling system from California Analytical Instruments, Inc. (model 602P) having a relative expanded uncertainty of $\pm$1~\% when compared to span gas volume fractions~\cite{Bundy:2007} was used to make gas concentration measurements. However, according to a study by Lock~et~al.~\cite{Lock:1}, the non-uniformity and movement of exhaust gases contribute to an estimated expanded uncertainty of $\pm$12~\%.

\subsubsection{Bi-Directional Probes}
Bi-directional probes with Setra 264 pressure transducers from Setra Systems, Inc. were used to measure gas velocity in doorways. Bryant~\cite{Bryant:FSJ2009} calculated an expanded uncertainty ranging from $\pm$14~\% to $\pm$22~\% for bi-directional probes of similar design.

\subsubsection{Heat Release Rate}
A positive displacement rotary gas meter was used to measure the volume flow rate of propane into the gas burners. The manufacturer, Romet Limited, reports a relative standard uncertainty of $\pm$2~\%~\cite{Romet:2014} for this type of meter (model RM-3000). A volumetric flow rate was calculated from the gas meter volume readings and used in conjunction with the heat of combustion of propane to calculate the heat release rate of the fire. The total expanded uncertainty for the heat release rate obtained from this method is estimated to be $\pm8$~\%.

@@

\chapter{Summary}
\label{chap:Summary}
Nine full-scale fire tests were conducted in two residential-sized structures to examine how changing ventilation patterns within a structure affects the fire environment. The fire source for each experiment was provided by a set of three diffusion flame burners with propane as the fuel. Various doors and vents were opened and closed during each test to change ventilation within the structure. A PPV fan was also used during some of the experiments to change the ventilation patterns. Local measurements of temperature, gas velocity, heat flux, and gas concentrations were collected at various locations throughout the structure during the experiments. The total volume of propane delivered to the burners was measured by a positive displacement rotary gas meter and was used to calculate the heat release rate of the fire during each test.
